### 2021-08-22 doc/thesis/chapters/conclusion-and-prospect.tex
Thesis chapters text. (#30)

This thesis covers a comprehensive variety of topics, that ranges from digital signal processing to system modeling and -design, IP design in different methods, verification, to hardware-software interaction in hardware.
The implementation of the FM radio receiver in multiple different methods, namely GNU Radio, Matlab, VHDL and HLS so that they achieve the same functionality, is a challenge, especially with the latter methods.
However, this challenge is solved within an end-to-end system design, that strongly supports the development process.\\

@@

The usage of Matlab as a tool to research DSP algorithms and strategies to demodulate a signal, in order to ultimately find a system model that can be implemented in hardware, turns out to be a powerful asset.
It allows the development in fast iterations, through short simulation times, to achieve a result quickly.
Based on this system model, the implementation of actual hardware IPs, with VHDL and HLS can be done more efficiently and in a direct way, because issues in the algorithm level are already fixed in the model at this point.
In combination with the simulation testbenches that share major parts, such as the input source, as well as a the analysis tool, the IP development works smoothly and efficiently.
The verification as a comparison against the Matlab system model gives instant feedback of whether the IP functions correctly.
Even if the system model changes, the testbench automatically adapts, since the input source files are re-generated by the model and thus picked up at the next iteration of the testbench simulation, so the hardware engineer can also adapt the IP code.
Once the IP is deployed on actual hardware, it is tested by the firmware, in order to check its correct functionality again, which closes the loop from the system model and the IP simulation to the hardware deployment.
This entire process proved itself to be a successful approach.\\

Regarding the implementation of the VHDL and HLS IPs and the accompanying comparison, the result turns out to be unexpected.
The initial expectation, that HLS is inefficient in terms of resource utilization, is contradicted.
Also the amount of time spent in the implementation and the metric of lines of code create a positive picture of HLS.
Further, the automatic generation of several files that are required to integrate the IP into a system design, such as interface definitions or software drivers, support this attitude.
VHDL may have its advantages in specific areas of application, but HLS is definitely a powerful tool that can achieve great results in less time.
This is especially true, when it comes to the development of complex functionalities, such as image- or video processing, with the support of built-in libraries like OpenCV.\\

However, the often heard statement that HLS enables any software developer to develop hardware IPs, can not be supported.
A developer that writes HLS code, still needs to have a strong hardware background, to be able to understand what is being generated in hardware.
It is important to know about the limitations of the target hardware, i.e. FPGA, in order to write an efficient design.
Therefore, the connection between specific HLS \cplusplus\ code lines, as well as the general structure and algorithm, and the hardware target need to be understood.\\

In prospect, there are many ideas on how the project can potentially be extended.
Additional implementation methods could be used to develop the FM radio receiver IP.
For example, Matlab can be used to create VHDL code, in a high-level synthesis fashion, which can be used to create the IP.
Regarding the verification environment, also the GNU Radio implementation could be included.
The common input that is produced by the Matlab model, and the common output analysis, which are both used for the VHDL and HLS testbenches, can be used in GNU Radio as well.
This would allow another comparison.
Looking at the HLS code, it could be evaluated how compiler-independent it is.
The question could be, if it is possible to write code that can be compiled with multiple compilers for different target platforms and vendors, e.g. Xilinx and Intel FPGAs, at the same time.
Furthermore, the various optimization strategies that Xilinx HLS provides could be evaluated on the existing HLS code.
Also, currently manually implemented functions, such as the FIR filter, could be replaced with build-in library functions.
Finally, the application software could be upgraded from the current bare-metal, real-time OS, to an embedded Linux OS which runs on the ARM CPU.
This would additionally allow the support of the RTL-SDR device to be used for live data streaming, instead of the current replay of the recorded antenna data from the SD card.\\

The amount of ideas for potential extensions to the project shows how comprehensive the project is and how many topics it covers.
Summing up, it can be said that the project revealed a range of interesting and partially unexpected insights into multiple topics, and that a lot of knowledge was gained during the elaboration of this thesis.

>>>>

This thesis covers a comprehensive variety of topics, including digital signal processing, system modeling and system design, IP design in different methods, verification, and hardware-software interaction in an actual hardware environment.
The implementation of the FM radio receiver in multiple different methods, namely GNU Radio, Matlab, VHDL and HLS, so that they all achieve the same functionality, is a challenge, especially with the latter methods.
However, this challenge is solved within an end-to-end system design, that strongly supports the development process.\\

@@

The usage of Matlab as a tool to research DSP algorithms and strategies to demodulate a signal, in order to ultimately find a system model that can be implemented in hardware, turns out to be a powerful asset.
Through short simulation times, it allows the development in fast iterations, which helps to achieve a result quickly.
Based on this system model, the implementation of actual hardware IPs, with VHDL and HLS can be done more efficiently and in a direct way, because issues in the algorithm level are already fixed in the model at this point.
In combination with the simulation testbenches, which share major parts, such as the input source, as well as the analysis tool, the IP development works smoothly and efficiently.
The verification as a comparison against the Matlab system model gives instant feedback of whether the IP functions correctly.
Even if the system model changes, the testbench automatically adapts, since the input source files are re-generated by the model.
These files are picked up at the next execution of the testbench simulation.
Consequently, the hardware engineer can also adapt the IP code accordingly.
Once the IP is deployed on actual hardware, it is tested by the firmware application, in order to check its correct functionality again, which closes the loop from the system model and the IP simulation to the hardware deployment.
This entire process proved itself to be a successful approach.\\

Regarding the implementation of the FM receiver in VHDL and HLS, and their subsequent comparison, the result turns out to be unexpected.
The initial expectation, that HLS may be inefficient in terms of resource utilization, is contradicted.
Also the relatively low amount of time that is required for the implementation, as well as the metric of lines of code, create a positive impression of HLS.
Further, the automatic generation of several files that are required to integrate the IP into a system design, such as interface definitions or software drivers, support this attitude.
VHDL may have its advantages in specific areas of application, but HLS is definitely a powerful tool that can achieve great results in less time.
This is especially true, when it comes to the development of complex functionalities, such as image- or video processing, with the support of built-in libraries like OpenCV.
Furthermore, the project presents a show-case, how HLS enables the concurrent development of a firmware application, which can already interface with the IP and get tested, before any actual hardware is available.\\

However, the often heard statement that HLS enables any software developer to develop hardware IPs, can not be supported.
A developer that writes HLS code, still needs to have a strong hardware background, to be able to understand what is being generated from HLS code in hardware.
It is important to know about the limitations of the target hardware, i.e. FPGA, in order to write an efficient design.
Therefore, the connection between specific HLS \cplusplus\ code lines, as well as the code's general structure and algorithm, and the hardware target need to be understood.\\

In prospect, there are many ideas on how the project can potentially be extended.
Additional implementation methods could be used to develop the FM radio receiver IP.
For example, Matlab can be used to create VHDL code, in a high-level synthesis fashion, which can be used to create the IP.
Regarding the verification environment, also the GNU Radio implementation could be included.
The common input that is produced by the Matlab model, and the common output analysis, which are both used for the VHDL and HLS testbenches, can be used in GNU Radio as well.
This would allow another comparison.
Looking at the HLS code, it could be evaluated how compiler-independent it is.
The question could be, if it is possible to write code that can be compiled with multiple HLS compilers for different target platforms and vendors, e.g. Xilinx and Intel FPGAs, at the same time.
Furthermore, the various optimization strategies that Xilinx HLS provides, could be evaluated on the existing HLS code.
Also, the currently manually implemented functions, such as the FIR filter, could be replaced with built-in library functions.
Finally, the application software could be upgraded from the current bare-metal, real-time OS, to an embedded Linux OS which runs on the ARM CPU.
This would additionally allow the support of the RTL-SDR device to be used for live data streaming, instead of the current replay of the recorded antenna data from the SD card.\\

The amount of ideas for potential extensions to the project shows how comprehensive the project is and how many topics it covers.
Summing up, it can be said, that the project revealed a range of very interesting and partially unexpected insights into multiple topics, and that a lot of knowledge was gained during the elaboration of this thesis.

### 2021-08-22 doc/thesis/chapters/deployment-on-hardware.tex
more of impl section
Thesis chapters text. (#30)

This chapter describes the additional parts of implementation that are necessary in order to run the implemented FM radio on actual hardware.

%%%%%%%%%%%%%%%%%%%%
\section{GNU Radio}

The GNU Radio implementations, both receiver and transmitter, are used in combination with external hardware.
The specific hardware that is used is briefly described here.\\

In order to interface with the real world, two different hardware devices are used.
Both devices are supported by GNU Radio and are briefly described here.

%%%%%%%%%%%%%%%%%%%%
\subsection{Ettus Research USRP b200mini}

The USRP b200mini is a device that can be used to create a software-defined radio.
It resembles the size of a business card and is developed by Ettus Research.
A signal range from 70~MHz to 6~GHz can be covered by the front end.
The board further features a Xilinx Spartan-6 FPGA that is user-programmable, which makes the device very flexible for various applications with the received signal \cite{USRPb200Mini}.\\

@@

The RTL-SDR is a cheap USB dongle, that was originally developed as a DVB-T TV tuner.
It is based on the RTL2832U chipset, which includes an 8-bit ADC and a digital signal processor.
The device delivers IQ data via the USB interface \cite{RTLSDR}.\\

@@

The ZedBoard is a development board that is developed by Xilinx.
It is specifically designed around the Xilinx Zynq-7000 SoC, which contains a dual-core ARM Cortex A9 processor and a programmable logic fabric, the FPGA.
The board further offers a range of features from external DDR3 memory, over various connectors, such as RJ45 Ethernet, USB, 3.5 mm audio, HDMI, Pmod, to buttons, LEDs and a display.\\

The most important features for this project are the audio line-out connector and the SD card port.
Additionally, the LEDs are used to display various status information during runtime of the FM receiver.

@@

The implemented version of the FM Radio Receiver on the ZedBoard is a working prototype at this point.
The implementation does not support live data streaming.
Instead, antenna data is recorded to a file with the RTL-SDR hardware, using a regular PC.
The antenna data is recorded as IQ data in the baseband.
The file is then stored onto an SD card, that is plugged into the ZedBoard.
The CPU on the SoC then reads it and replays the antenna data to the FPGA.
In the future, this design may be extended to use an embedded Linux operating system.
In that way, the RTL-SDR could be connected to the ZedBoard PCB directly via USB, which would enable a live data stream from the antenna to the FPGA.
However, the chosen solution is sufficient for the aims of this thesis.

@@

The block diagram in Figure~\ref{fig:bd_impl_vivado} shows the implemented architecture that is used in the SoC.
The various design domains in and around the SoC are depicted in different visual representations.
The FPGA domain blocks are highlighted in gray, with the main IPs being surrounded by emphasized lines.
The processor system domain is shown as white blocks.
Blocks that are external to the SoC are drawn with dashed lines.

@@

The CPU is the main actor that determines any actions on the system.
First, it reads a file from the external SD card and stores this data in a dedicated area in the memory.
Next, the DMA is configured and its data transfer mode is initiated, so that the DMA starts to transfer data from the mentioned memory area towards the FPGA.
The DMA is configured in a mode that creates a continuous transfer, by looping through the data -- once the end is reached, the transfer starts at the beginning of the memory area.
Any data connection in the following processing chain is implemented as an AXI stream bus.
The first block after the DMA is an AXI Stream Switch.
It can switch its single input to be forwarded to either of the outputs.
The switch is configured by the CPU, to enable one output at a time.
Consequently, one of the FM radio receiver IPs is provided with data.
The IP processes this data to produce an audio output, which is delivered to another AXI Stream Switch.
It fulfills the function to either select the output of above or below IP, in coordination with the first switch, obviously.
The next block is an AXI Stream Broadcaster, which takes a single stream input and simply duplicates it at its output, so that the outputs are an exact copy of the input stream.
One of these duplicates is sourced to an AXI-Stream-To-I2S converter, which generates the I2S signals for the external audio codec chip.
The second duplicate is connected to a FIFO, which simply stores the generated output data.
Once the FIFO is almost full, it asserts an interrupt to the CPU, which triggers it to read out the entire FIFO buffer.
The CPU then stores this data on the external SD card again.
In that way, the loop of data is closed, which enables the verification of the respective IPs' correct functionality.
Therefore, this system design can be described as a hardware-in-the-loop system, in the point of view of the FPGA IPs.
This ability refers to the knowledge gained with the SIMILAR concept, that is introduced in Section~\ref{sec:SimilarProcess}.
Specifically, this implemented system design enables the application of the steps \textit{Assess Performance} and \textit{Re-evaluate}, since the IPs' performance can be measured and their design can be re-evaluated.\\

@@

The application software is running on a single core of the dual-core ARM Cortex A9 CPU and is built upon a real-time operating system, namely FreeRTOS.
Xilinx provides a ported version of this RTOS with the installation of their software development kit, the Xilinx SDK.
The SDK also includes drivers to support various functionalities, such as support for accessing a FAT file system on the external SD card.\\

The respective software drivers for several IPs that are used in the FPGA design are automatically imported into the SDK project.
They are embedded in the so-called hardware definition file (HDF) that is generated by Vivado after the synthesis process, and provided together with the bitstream file.
It is a file with the extension \textit{*.hdf} and can be opened like a regular archive file, with any ZIP archive viewer.\\

Generally, the application software for this project is developed as an object-oriented \cplusplus\ program.
It uses common software development patterns and a class structure that fits the purpose.
The main tasks of the application are to configure several IPs in the FPGA, so that they can fulfill their respective purpose in the system, and to read the recorded data from the SD card, which is then streamed to the FPGA.
It also provides a basic user interface over the serial console, with which a user can interact with the FM Radio Receiver product.\\

However, software design is not the main focus of this thesis, and thus no further details are discussed here.

>>>>

This chapter describes the additional parts of implementation that are necessary, in order to run the implemented FM radio receiver on actual hardware.

%%%%%%%%%%%%%%%%%%%%
\section{GNU Radio}
\label{sec:deployment-on-hw:gnu-radio-devices}

The GNU Radio implementations, both receiver and transmitter, are used in combination with external hardware.
Two different hardware devices are used to interface with the real world, which are both supported by GNU Radio directly.
The specific devices are briefly described here.

%%%%%%%%%%%%%%%%%%%%
\subsection{Ettus Research USRP B200mini}

The USRP B200mini is a device that can be used to create a software-defined radio.
It resembles the size of a business card and is developed by Ettus Research.
A signal range from 70~MHz to 6~GHz can be covered by the front end.
The board further features a Xilinx Spartan-6 FPGA that is user-programmable, which makes the device very flexible for various applications with the received signal \cite{USRPb200Mini}.\\

@@

The RTL-SDR is a low-cost USB dongle, that was originally developed as a DVB-T TV tuner.
It is based on the RTL2832U chipset, which includes an 8-bit ADC and a digital signal processor.
The device delivers IQ data via the USB interface \cite{RTLSDR}.\\

@@

The ZedBoard is a development board that is developed by Xilinx.
It is specifically designed around the Xilinx Zynq-7000 SoC, which contains a dual-core ARM Cortex~A9 processor and a programmable logic fabric, the FPGA.
The board further offers a range of features like external DDR3 memory, various connectors, such as RJ45 ethernet, USB, 3.5~mm audio in- and output, HDMI, Pmod, as well as user interface hardware such as buttons, LEDs and a display.\\

The most important features for this project are the 3.5~mm audio line-out connector and the SD card port.
Additionally, the LEDs are used to display various status information during the runtime of the FM receiver.

@@

The implemented version of the FM radio receiver on the ZedBoard is a working prototype at this point.
The implementation does not support live data streaming.
Instead, antenna data is recorded to a file with the RTL-SDR hardware, using a regular PC.
The data is recorded as IQ data in the baseband.
The file is then stored onto an SD card, that is plugged into the ZedBoard.
The CPU on the SoC reads it and replays the antenna data to the FPGA.
In the future, this design may be extended to use an embedded Linux operating system.
In that way, the RTL-SDR could be connected to the ZedBoard PCB directly via USB, which would enable a live data stream from the antenna to the FPGA.
However, the chosen solution is sufficient for the aims of this thesis.

@@

The block diagram in Figure~\ref{fig:bd_impl_vivado} shows the implemented architecture that is used in the SoC.
The various design domains in, and around the SoC, are depicted in different visual representations.
The FPGA domain blocks are highlighted in gray, with the main IPs being surrounded by emphasized lines.
The processor system domain is shown as white blocks.
Blocks, which are external to the SoC, are drawn with dashed lines.

@@

The CPU is the main actor that determines any actions on the system.
First, it reads a file from the external SD card and stores this data in a dedicated area in the memory.
Next, the DMA is configured and its data transfer mode is initiated, so that the DMA starts to transfer data from the mentioned memory area towards the FPGA.
The DMA is configured in a mode that creates a continuous transfer, by looping through the data -- once the end is reached, the transfer starts at the beginning of the memory area.
Any data connection in the following processing chain is implemented as an AXI stream bus.
The first block after the DMA is an AXI Stream Switch.
It can switch its single input to be forwarded to either of the outputs.
The switch is configured by the CPU, to enable one output at a time.
Consequently, one of the FM radio receiver IPs is provided with data.
The IP processes this data to produce an audio output, which is delivered to another AXI Stream Switch.
It fulfills the function to either select the output of above or below IP, in coordination with the first switch, obviously.
The next block is an AXI Stream Broadcaster, which takes a single stream input and simply duplicates it at its output, so that the outputs are an exact copy of the input stream.
One of these duplicates is sourced to an AXI-Stream-To-I2S Converter, which generates the I2S signals for the external audio codec chip.
The second duplicate is connected to a FIFO, which simply stores the generated output data.
Once the FIFO is almost full, it asserts an interrupt to the CPU, which triggers it to read out the entire FIFO buffer.
The CPU then stores this data on the external SD card.
In that way, the loop of data is closed, which enables the verification of the respective IP's correct functionality.
Therefore, this system design can be described as a hardware-in-the-loop system, in the point of view of the FPGA IPs.
This ability refers to the knowledge gained with the SIMILAR concept, that is introduced in Section~\ref{sec:SimilarProcess}.
Specifically, this implemented system design enables the application of the steps \textit{Assess Performance} and \textit{Re-evaluate}, since the IPs' performance can be measured and their design can be re-evaluated.\\

@@

The application software is running on a single core of the dual-core ARM Cortex~A9 CPU and is built upon a real-time operating system, namely FreeRTOS.
Xilinx provides a ported version of this RTOS with the installation of their software development kit, the Xilinx SDK.
The SDK also includes drivers to support various functionalities, such as support for accessing a FAT file system on the external SD card.\\

The respective software drivers for several IPs, which are used in the FPGA design, are automatically imported into the SDK project.
They are embedded in the so-called hardware definition file (HDF) that is generated by Vivado after the synthesis process, and is provided together with the bitstream file.
It is a file with the extension \textit{.hdf} and can be opened like a regular archive file, with any archive viewer.\\

Generally, the application software for this project is developed as an object-oriented \cplusplus\ program.
It uses common software development patterns and a class structure that fits the purpose.
The main tasks of the application are to configure several IPs in the FPGA, so that they can fulfill their respective purpose in the system, and to read the recorded data from the SD card, which is then streamed to the FPGA.
It also provides a basic user interface over the serial console, which a user can use to interact with the FM Radio Receiver product.
However, software design is not the main focus of this thesis, and thus, no further details are discussed here.

### 2021-08-22 doc/thesis/chapters/hls.tex
Thesis chapters text. (#30)

\begin{itemize}
  \item \textbf{Create input vectors}\\
  In this first step, the above explained, regular C simulation is executed.
  Hereby, the inputs that are sent to the DUT are stored as \textit{input vectors}.
  \item \textbf{Perform RTL simulation}\\
  Next, the previously generated input vectors are applied to the DUT.
  At this point, the DUT is already translated into RTL code, so the code that is used here is either VHDL or Verilog code.
  The default simulator is the Vivado Simulator (XSim).
  However, various other simulators like ModelSim are supported.
  The generated outputs of this step are stored as \textit{output vectors}.
  \item \textbf{Verify the results}\\
  The final verification step sources the output vectors back to the C testbench, which compares the outputs with the expected results.\\
\end{itemize}

\noindent
This type of simulation takes a significantly longer time to execute, because timing information now actually needs to be considered for the RTL code, which is done by using an HDL simulator.

>>>>

\begin{itemize}
  \item \textbf{Create input vectors}\\
  In this first step, the above explained, regular C simulation is executed.
  Hereby, the inputs that are sent to the DUT are stored as \textit{input vectors}.\\

  \item \textbf{Perform RTL simulation}\\
  Next, the previously generated input vectors are applied to the DUT.
  At this point, the DUT is already translated into RTL code, so the code that is used here is either VHDL or Verilog code.
  The default simulator is the Vivado Simulator (XSim).
  However, various other simulators like ModelSim are supported.
  The generated outputs of this step are stored as \textit{output vectors}.\\

  \item \textbf{Verify the results}\\
  The final verification step sources the output vectors back to the C testbench, which compares the outputs with the expected results.\\
\end{itemize}

\noindent
The execution of this type of simulation takes significantly longer, because timing information now actually needs to be considered for the RTL code, which is done by using an HDL simulator.

### 2021-08-22 doc/thesis/chapters/results.tex
Thesis chapters text. (#30)

The system design architecture that is chosen in this project covers several parts of the system.
This includes a reference model, the development of the IP itself, including the simulation testbench, as well as the final integration into the productive system.
Because of this modular and structured approach, that is given by this system design, it can be applied to almost any DSP project.
The main processing IP can simply be replaced with another implementation, while the environment does not need to be modified.
The reason therefor is, that DSP projects almost always use streaming interfaces for their inputs and outputs.
Also a configuration interface is commonly used, to adapt certain parameters of the IP during runtime.

@@

The aim is to create IP blocks, that can be integrated into a block design in the Xilinx Vivado software.
The IP blocks should be able to be connected with their AXI interfaces, in order to successfully communicate and interact with the embedded hard-processor, the ARM CPU.
In the case of HLS, several files that are required for this integration are generated automatically.
The files are already placed into the specific folder structure, that Vivado requires to recognize it as an IP block.
In the case of VHDL on the other hand, this needs to be done manually.
There is an assistant, the IP Packager, which helps the developer to create the IP.
However, at this point the folder structure already needs to be in place, so that the assistant works correctly, which again requires manual action.
Regarding the top-level interfaces, the IP Packager supports an auto-detection of the signals that belong to the respective interface types, as long as they follow the expected naming guidelines.
However, the interface itself needs to be defined and updated manually in the IP Packager.
This is especially important, if the interfaces change in the VHDL code.\\

However, also the auto-generation of the HLS IP has characteristics that require special attention.
One issue during the implementation was, that the auto-generated software driver did not get imported into the Vivado SDK.
The reason therefor was, that the top-level function of the IP did not have the exact same naming as the Vivado HLS project.
For issues like that, the Eclipse-based SDK does not give any error output or any kind of information.
This however is a general issue with the SDK, where error handling is poor and error log outputs are not existent, or at least not useful to the developer, so that the SDK sometimes crashes without an obvious reason.
It is to note, that version v2018.2 is used for the entire Vivado toolchain, which is not the latest version that is currently available.
Thus, some of the occuring issues may already be fixed in the newer versions.\\

However, both IPs could successfully be integrated into the Vivado block design, and the tests on the hardware show that they function correctly within the system.

@@

The hardware utilization of the HLS and VHDL implementations are compared in detail in this section.
The numbers are taken from the final utilization report, that is created by Vivado at the end of the bitstream creation, since this is assumed to be the most accurate report that shows the actual utilization.\\

Figure~\ref{fig:hardware_utilization} shows the utilization numbers of the different FPGA resources.
It reveals that the manual VHDL implementation clearly requires more resources in the FPGA.
The study of the detailed utilization report however does not enable a deeper direct comparison of the respective single blocks, such as VHDL FIR filter against HLS FIR filter.
This is because the HLS compiler automatically applies optimizations, so that the direct connection between code and resource report is partially lost.
Examples for optimizations are inlining of functions, re-ordering of operations and time-sharing.
The latter has an especially large impact on the resource utilization.
Generally, in hardware design there is a trade-off between resource utilization and throughput, which leads to time-sharing.
For example, an entirely parallel-duplicated pipeline can process the double amount of data, but also requires the double amount of resources.
On the other hand, time-sharing a resource may take the double amount of time, thus only processes half of the data, but also only requires half of the resource utilization.
In HLS, this trade-off can be influenced and determined by the developer.
Depending on the requirements of the design, the optimization may either be set towards a lower resource utilization, or towards a high throughput.
In the current HLS implementation, no measures are taken towards any optimization and thus, the compiler is free to decide.
However, the only requirement is the correct functionality within the given situation, of course.
This specifically means that the throughput needs to be high enough, in order to process the audio signal output accordingly.\\

Nevertheless, the resource report allows the comparison of the configuration interface, which is the AXI4-Lite interface.
In VHDL, it requires 44 LUTs and 54 FFs, while HLS consumes only 36 LUTs and 37 FFs.
So it seems that the manual VHDL implementation requires more resources in all the comparable parts.
By further investigating the report, the VHDL version apparently utilizes a large amount of LUTs for the delays, which are used to clock-align the FIR output signals.
For an optimization of the VHDL design, this would be a good point to start at.
However, for this project, both implementation variants are compared without any optimizations from the developer's side.

@@

@@

Latency is the time it takes for the design to produce an output sample.
The latency of the HLS version is taken from the report that is produced after the C-synthesis of the HLS compiler.
The report shows an estimated resource utilization, as well as the latency as a number of clock cycles.
The latency is given with 2712 clock cycles.
For the VHDL version, a similar report is not available.
Instead, the testbench of the integration test is executed, which produces a signal wave form.
The latency is measured from this wave form and results in a number of 2470 clock cycles.
With regards to latency, the two implementation variants appear to have a similar behaviour.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{GNU Radio}

This section is specifically (TODO)

>>>>

The system design architecture, that is chosen in this project, covers several parts of the system.
This includes a reference model, the development of the IP itself, including the simulation testbench, as well as the final integration into the productive system.
Because of this modular and structured approach, given by this system design, it can be applied to almost any DSP project.
The main processing IP can simply be replaced with another implementation, while the environment does not need to be modified.
The reason therefor is, that DSP projects almost always use streaming interfaces for their inputs and outputs.
Also, a configuration interface is commonly used, to adapt certain parameters of the IP during runtime.

@@

The aim is to create IP blocks, that can be integrated into a block design in the Xilinx Vivado software.
These IP blocks should be able to be connected with their AXI interfaces, in order to successfully communicate and interact with the embedded hard-processor, the ARM CPU, as well as other IPs that exist in the design.
In the case of HLS, several files that are required for this integration are generated automatically.
The files are already placed into the specific folder structure, that Vivado requires to recognize it as an IP block.
In the case of VHDL on the other hand, this needs to be done manually.
There is an assistant, the IP Packager, which helps the developer to create the IP.
However, at this point, the folder structure already needs to be in place, so that the assistant works correctly, which again requires manual action.
Regarding the top-level interfaces, the IP Packager supports an auto-detection of the signals that belong to the respective interface types, as long as they follow the expected naming guidelines.
However, the interface itself needs to be defined and updated manually in the IP Packager.
This is especially important, if the interfaces change in the VHDL code.\\

However, also the auto-generation of the HLS IP has characteristics that require special attention.
One issue during the implementation was, that the auto-generated software driver did not get imported into the Vivado SDK.
The reason therefor was, that the top-level function of the IP did not have the exact same naming as the Vivado HLS project.
This is a non-documented requirement, and furthermore, for issues like that, the Eclipse-based SDK does not give any error output or any kind of meaningful information.
This however is a general issue with the SDK, where error handling is poor and error log outputs are not existent, or at least not useful to the developer, so that the SDK sometimes crashes without an obvious reason.
It is to note, that version v2018.2 is used for the entire Vivado toolchain, which is not the latest version that is currently available.
Thus, some of the occuring issues may already be fixed in the newer versions.\\

However, both IPs could successfully be integrated into the Vivado block design in the end, and the tests on the hardware show that they function correctly within the system.

@@

The hardware utilization of the HLS and VHDL implementation is compared in detail in this section.
The numbers are taken from the final utilization report, that is created by Vivado at the end of the bitstream creation, i.e. post-synthesis, since this is assumed to be the most accurate report, which shows the actual utilization, as it is being deployed to hardware.

Figure~\ref{fig:hardware_utilization} shows the utilization numbers of the different FPGA resources.
It reveals, that the manual VHDL implementation clearly requires more resources in the FPGA.
The study of the detailed utilization report, however, does not enable a deeper direct comparison of the respective single blocks, such as VHDL FIR filter versus HLS FIR filter.
This is because the HLS compiler performs some kind of name mangling, which re-names functions, and automatically applies optimizations, so that the direct connection between code and resource report is partially lost.
Examples for optimizations are inlining of functions, re-ordering of operations and time-sharing.
The latter has an especially large impact on the resource utilization.
Generally in hardware design, there is always a trade-off between resource utilization and throughput, which ultimately leads to time-sharing.
For example, an entirely parallel-duplicated pipeline can process the double amount of data, but also requires the double amount of resources.
On the other hand, time-sharing a resource may take the double amount of time, thus only processes half of the data, but also only requires half of the resource utilization.
In HLS, this trade-off can be influenced and determined by the developer, by the usage of optimization directives.
Depending on the requirements of the design, the optimization may either be set towards a low resource utilization, or towards a high throughput.
In the current HLS implementation, no measures are taken towards any optimization and thus, the compiler is free to decide.
However, the only requirement is the correct functionality within the given situation, of course.
This specifically means that the throughput needs to be high enough, in order to process the audio signal output accordingly.\\

@@

Nevertheless, the resource report allows the direct comparison of one specific unit, the configuration interface, which is the AXI4-Lite interface.
In VHDL, it requires 44~LUTs and 54~FFs, while HLS only consumes 36~LUTs and 37~FFs.
It seems, that the manual VHDL implementation requires more resources in all the comparable parts.
By further investigating the report, the VHDL version apparently utilizes a large amount of LUTs for the delays, which are used to clock-align the FIR output signals.
For an optimization of the VHDL design, this would be a good point to start at.
However, for this project, both implementation variants are compared without any optimizations from the developer's side.

@@

Latency is the time it takes for the design to produce an output sample.
The latency of the HLS version is taken from the report that is produced after the C-synthesis of the HLS compiler.
The report shows an estimated resource utilization, as well as the latency as a number of clock cycles.
The latency is given with 2712~clock~cycles.
For the VHDL version, a similar report is not available.
Instead, the testbench of the integration test is executed, which produces a signal wave form.
The latency is measured from this wave form and results in a number of 2470~clock~cycles.
With regards to latency, the two implementation variants appear to have a similar behaviour.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\subsection{GNU Radio}

%This section is specifically (TODO)
% - show that the transmitter works (show photo of radio display in my car)
% - show that the receiver works (screenshot of decoded RDS data, etc.)

### 2021-08-22 doc/thesis/front/kurzfassung.tex
Thesis chapters text. (#30)

Diese Masterarbeit wurde als Abschlussarbeit im Rahmen des Master Studiengangs Embedded Systems Design an der Fachhochschule Oberösterreich, Campus Hagenberg ausgearbeitet.
Begleitend zu den theoretischen Themen dieser Arbeit wurde ein umfangreiches, praktisches Projekt entwickelt.\\

Das Hauptthema dieser Arbeit ist es, ein System-Design für ein Embedded System zu entwickeln, welches alles Wissen über sämtliche Disziplinen und Themen umfasst, die im Rahmen des Studiums behandelt wurden.
Mit einem Entwurfsansatz auf System-Ebene wird in mehreren Implementierungsvarianten ein FM Radio Empfänger entwickelt.
Zielarchitekture ist ein System-on-Chip, das aus FPGA und CPU besteht.\\

Im Speziellen werden zwei Varianten für die Implementierung eingesetzt -- High-Level Synthese und die manuelle Implementierung von VHDL code.
Das betrifft hauptsächlich die digitale Signalverarbeitungskette, sowie mehrere Schnittstellen, anhand welcher der Hardware Entwurf mit seiner Umgebung kommunizieren kann.
Die beiden Varianten werden anhand verschiedener Kennzahlen verglichen.
Beispielsweise werden hierfür die Zeitdauer der Implementierung, die Anzahl der Programmzeilen, die Effizienz in der Nutzung der Hardware, sowie der generelle Eindruck bezüglich der Benutzbarkeit während der Entwicklung, herangezogen.
Am Ende wird ein produktives System auf echter Hardware erstellt, inklusive einer Firmware Anwendung, um die Interaktion mit einem Benutzer zu ermöglichen.\\

Der gesamte Entwickungsprozess wird immer in dem Gesichtspunkt des Systementwurfs betrachtet.
Deshalb wird im ersten Schritt ein Systemmodell mit Matlab entwickelt.
Dieses erlaubt es, Algorithmen für die digitale Signalverarbeitung, und generell ein System zu erarbeiten, welches später in Hardware implementiert werden kann.
Weiters wird GNU Radio dazu verwendet, um einen Prototypen zu entwickeln, der bereits auf echter Hardware lauffähig ist.
Das gesamte System, inklusive der beiden Implementierungsvarianten und dem Matlab Systemmodell, verwendet eine gemeinsame Strategie und Umgebung zur Verifikation der einwandfreien Funktion des Hardwareentwurfs.
Diese Verifikation wird in Simulation, als auch direkt in der Hardware durchgeführt, wodurch sich der Kreis des Ansatzes als Systementwurf schließt.

>>>>

Diese Masterarbeit wurde als Abschlussarbeit im Rahmen des Masterstudiengangs Embedded Systems Design an der Fachhochschule Oberösterreich, Campus Hagenberg ausgearbeitet.
Begleitend zu den theoretischen Themen dieser Arbeit wurde ein umfangreiches, praktisches Projekt entwickelt.\\

Das Hauptthema dieser Arbeit ist es, ein System-Design für ein Embedded System zu entwickeln, welches sämtliches Wissen über alle Disziplinen und Themen umfasst, die im Rahmen des Studiums behandelt wurden.
Mit einem Entwurfsansatz auf System-Ebene wird in mehreren Implementierungsvarianten ein FM Radio Empfänger entwickelt.
Zielarchitektur ist ein System-on-Chip, das aus FPGA und CPU besteht.\\

Im Speziellen werden zwei Varianten für die Implementierung eingesetzt -- High-Level Synthese und die manuelle Implementierung von VHDL Code.
Dies betrifft hauptsächlich die digitale Signalverarbeitungskette, sowie mehrere Schnittstellen, anhand welcher der Hardwareentwurf mit seiner Umgebung kommunizieren kann.
Die beiden Varianten werden anhand verschiedener Kennzahlen verglichen.
Beispielsweise werden hierfür die Zeitdauer der Implementierung, die Anzahl der Programmzeilen, die Effizienz in der Nutzung der Hardware, sowie der generelle Eindruck bezüglich der Benutzbarkeit während der Entwicklung, herangezogen.
Am Ende wird ein produktives System auf echter Hardware erstellt, inklusive einer Firmware Anwendung, um die Interaktion mit einem Benutzer zu ermöglichen.\\

Der gesamte Entwickungsprozess wird immer in dem Gesichtspunkt des Systementwurfs betrachtet.
Deshalb wird im ersten Schritt ein Systemmodell mit Matlab entwickelt.
Dieses erlaubt es, Algorithmen für die digitale Signalverarbeitung, und generell ein Gesamtsystem, zu erarbeiten, welches später in Hardware implementiert werden kann.
Weiters wird GNU Radio dazu verwendet, um einen Prototypen zu entwickeln, der bereits auf echter Hardware lauffähig ist.
Das gesamte System, inklusive der beiden Implementierungsvarianten und dem Matlab Systemmodell, verwendet eine gemeinsame Strategie und Umgebung zur Verifikation der einwandfreien Funktion des Hardwareentwurfs.
Diese Verifikation wird in Simulation, als auch direkt in der Hardware durchgeführt, wodurch sich der Kreis des Ansatzes als Systementwurf schließt.
